Project 1 Reproducible research Coursera
========================================================

## Load data and add an extra column for the day

```{r}
data<-read.csv(file='activity.csv',stringsAsFactors=FALSE,colClasses=c('numeric','Date','numeric'))
data<-within(data,day<-as.numeric(date-min(date)))
```

## What is mean total number of steps taken per day?

```{r}
library(data.table)
data<-data.table(data)
data_plot<-data[,sum(steps,na.rm=TRUE),by=day]
barplot(data_plot$V1, names.arg=data_plot$day,ylim=c(0,max(data_plot$V1)),main='Nb steps per day',xlab='days',ylab='Nb of steps')
summary(data_plot$V1)[c('Mean','Median')]
```

## What is the average daily activity pattern?

```{r}
data_plot<-data[,mean(steps,na.rm=TRUE),by=interval]
within(data_plot,plot(interval,V1,type="l",main='Average nb steps per interval',xlab='Daily time interval',ylab='Nb steps'))
data_plot[which.max(data_plot$V1)]
```

## Imputing missing values
Complete missing data with the mean steps of 5-minute interval calculated previously
```{r}
sum(!complete.cases(data))
index_for_missing_values<-which(!complete.cases(data))
completed_data<-data
filling_values<-data_plot
for(i in 1:length(index_for_missing_values)) {
  current_interval<-data[index_for_missing_values[i],interval]
  completed_data[index_for_missing_values[i],steps:=filling_values[interval==current_interval,V1]]
}

data_plot<-completed_data[,sum(steps,na.rm=TRUE),by=day]
barplot(data_plot$V1, names.arg=data_plot$day,ylim=c(0,max(data_plot$V1)),main='Nb steps per day',xlab='days',ylab='Nb of steps')
summary(data_plot$V1)[c('Mean','Median')]
```

## Are there differences in activity patterns between weekdays and weekends?

```{r}
data<-completed_data
column_to_add<-weekdays(data$date,abbreviate=TRUE)
column_to_add<-sub("^Mon|^Tue|^Wed|^Thu|^Fri","weekday",column_to_add,perl=TRUE)
column_to_add<-sub("^Sat|^Sun","weekend",column_to_add,perl=TRUE)
data$wk_pattern<-as.factor(column_to_add)
data_plot<-data[,mean(steps,na.rm=TRUE),by=list(interval,wk_pattern)]
par(mfrow=c(2,1))
plot(V1 ~ interval,col=wk_pattern,data=data_plot[wk_pattern=='weekend',],type='l',main='weekend',ylab='Number of steps',xlab='Interval')
plot(V1 ~ interval,col=wk_pattern,data=data_plot[wk_pattern=='weekday',],type='l',main='weekday',ylab='Number of steps',xlab='Interval')

